---
title: "the interaction between foehn and temperature"
author: "Tino Schneidewind"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r data, echo=FALSE, warning=FALSE, message=FALSE}
### PACKAGES ####
library(dlnm);library(splines);library(ggplot2);library(viridis);library(gnm); library(dplyr);library(knitr);library(lmtest) library(ggstance)

data = read.csv()

# index to include only stratum that have hospitalization counts
data$stratum_dow = as.factor(data$stratum_dow)
ind_dow = tapply(data$all, data$stratum_dow, sum)


# define the maximum lag distance we account for
maxlago <- 3

# mmt function
source("/functions/findmin.R")

# crossbasis temperature
cb.temp <- crossbasis(data$temperature,
                      lag=21,
                      argvar=list(fun="ns", knots = quantile(data$temperature, c(.5,.9), na.rm=TRUE)),
                      arglag=list(fun="ns", knots = logknots(21,3)),
                      group = data$station)

# binary foehn
foehn_bin     <- ifelse(data$foehn_wind >= 72, 0, 1)
foehn_bin_rev <- ifelse(foehn_bin == 1, 0, 1)

# modifier functions
modif     <- cb.temp * foehn_bin
modif_rev <- cb.temp * foehn_bin_rev

# groups
groups_id = colnames(data)[c(3,9,10,24,25, 13, 14, 11, 15, 12)]
letters = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j")


# colors
colors <- viridis(3, option = "viridis")
foehn_col = colors[2]
temp_col = colors[3]

```

<br>

This Markdown document presents and summarizes the findings of an analysis concerning the following research question.

<br>

#### *Results for RQ2+RQ3: is the temperature-hospitalizations association significantly modified by foehn winds in Switzerland? are certain subgroups more affected than others?*

<br>

```{r results_all, echo=FALSE, fig.width=3.3, fig.height=3.3, fig.align='center'}
foehn_col = "green4"
temp_col = "gold2"


for (i in 1:length(groups_id[1])) {

  colvar  =  groups_id[i]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  mod_modif_rev <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20)
  pred_modif2  <- crosspred(cb.temp, mod_modif_rev, cen = 20)
  
  # get min value of both predictions and use for centering
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))

  # predict with new min values
  pred_modif_new     <- crosspred(cb.temp, mod_modif, cen = min1, by = .1)
  pred_modif_rev_new <- crosspred(cb.temp, mod_modif_rev, cen = min1, by = .1)
  
  plot(pred_modif_new,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)

  
  lines(pred_modif_rev_new,           ## cumulative exposure
     "overall",
     col = temp_col,
     ci = "area",
     ci.arg = list(col = alpha(colour = temp_col, .15)),
     lwd = 2)
  
  abline(v =quantile(data$temp, .99), col = "black", lty = 2)
  
    legend("top", ncol = 1, legend = c("temp with interaction", "temp without interaction"), col = c("green4", "gold2"),
         bty = "n", lwd=c(2,2), cex = 0.7)
  
  text(25, 2.5 , labels = "(a)", pos = 4, cex = 0.7)
  }

```

*Figure 3b: Cumulative relative risk from temperature exposure (Model 4) for all-cause hospitalizations divided into foehn and non-foehn days with 95% confidence intervals. The dotted line shows the temperature that corresponds to the 99th percentile of the temperature distribution (24.7 °C).*


```{r results_subgroups_lollipop_figure, echo=FALSE}
# interaction between foehn winds and temperature in subgroups at 24.7°C

table_estimates = data.frame(categories = c("all","all", "mal", "mal","fem","fem","y64","y64","o64","o64",
                                            "cvd","cvd","resp","resp","inf","inf","uri","uri","ment","ment"),
                             model = rep(c("temp + foehn", "temp - foehn"), 10),
                             pred = rep(NA, 20),
                             CI_low = rep(NA, 20),
                             CI_high = rep(NA, 20))

foehn_col = "green4"
temp_col = "gold2"


for (i in 1:nrow(table_estimates)) {

  colvar  =  table_estimates$categories[i]
  
  if( i %% 2 != 0) {
    formula1 = as.formula(paste0(colvar, "~ cb.temp + modif"))
  } else {
    formula1 = as.formula(paste0(colvar, "~ cb.temp + modif_rev")) 
  }
    
  # calculate minimum hospitalization temperature for each subpopulation
  formulamin = as.formula(paste0(colvar, "~ cb.temp + modif"))
  mod_modif     <-gnm(formulamin, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20, cumul=FALSE)
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))
  
  
  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # predict with new min values
  pred_modif_new  <- crosspred(cb.temp, mod_modif, cen = min1, cumul=FALSE, by = .1)
  
  
  table_estimates$pred[i]    = pred_modif_new$allRRfit["24.7"]
  table_estimates$CI_low[i]  = pred_modif_new$allRRlow["24.7"]
  table_estimates$CI_high[i] = pred_modif_new$allRRhigh["24.7"]


}

table_estimates$categories <- factor(table_estimates$categories, 
                                     levels = rev(unique(table_estimates$categories)))

table_estimates$model <- factor(table_estimates$model, 
                                levels = c("temp + foehn", "temp - foehn"))



ggplot(table_estimates, aes(y = categories, x = pred, color = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high), 
                  position = position_dodgev(height = 0.4)) +
  labs(title = "",
       x = "relative risk", y = "") +
  theme_minimal() +
  scale_color_manual(
    values = c("temp + foehn" = foehn_col, "temp - foehn" = temp_col), 
    labels = c("temp + foehn" = "Model 4 on foehn days", "temp - foehn" = "Model 4 on non-foehn days"),  # Map the levels to custom labels
    name = "model:"  # Custom legend title
  ) +
  annotate("text", x = Inf, y = Inf, label = "(b)", 
           hjust = 1.2, vjust = 1.2, size = 3) +
  
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 8),  
    axis.text.y = element_text(size = 8),    
    axis.title.x = element_text(size = 10),  
    plot.margin = unit(c(0, 1, 0, 0), "cm"), 
    axis.line = element_line(color = "black"), 
    panel.grid.minor = element_blank(),      
    panel.grid.major.y = element_blank(),  
    panel.grid.major.x = element_blank(),  
    legend.margin = margin(.5, .5, .5, .5),
    legend.background = element_rect(fill = "white", color = "white"),
  ) +
  guides(color = guide_legend(ncol = 1))
```

*Figure 3c: Cumulative relative risk (Model 4) for subgroups at 24.7 °C with 95% confidence intervals. *



```{r results_all_interaction, echo=FALSE, fig.width=3.3, fig.height=3.3, fig.align='center'}
# all cause analysis interaction plot

foehn_col = "green4"
temp_col = "gold2"


for (i in 1:length(groups_id[1])) {

  colvar  =  groups_id[i]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <- gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction on the modifier
  pred_modif   <- crosspred(modif_rev, mod_modif, cen = 15.5)
  
  plot(pred_modif,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)
  
  }

```

*sFigure 5: Interaction between foehn and temperature with 95% confidence interval for all-cause hospitalizations. *


## Supplementary Material

```{r results_subgroups, echo=FALSE, fig.width=9, fig.height=9, fig.align='center'}
# subgroups
for (i in 1:length(groups_id[2:10])) {

  colvar  =  groups_id[i+1]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  mod_modif_rev <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20, cumul=FALSE)
  pred_modif2  <- crosspred(cb.temp, mod_modif_rev, cen = 20, cumul=FALSE)

  # get min value of both predictions and use for centering
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))

  # predict with new min values
  pred_modif_new     <- crosspred(cb.temp, mod_modif, cen = min1, cumul=FALSE, by = .1)
  pred_modif_rev_new <- crosspred(cb.temp, mod_modif_rev, cen = min1, cumul=FALSE, by = .1)
  
  plot(pred_modif_new,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .2)),
     lwd = 2,
     main ="",
     ylim = c(0.7,3)
     )

  
  lines(pred_modif_rev_new,           ## cumulative exposure
     "overall",
     col = temp_col,
     ci = "area",
     ci.arg = list(col = alpha(colour = temp_col, .2)),
     lwd = 2)
  
  text(25, 3, labels = paste0("(", letters[i], ")"), pos = 2)}

```

*sFigure 6: Cumulative response ratios for (a) all-cause, (b) male, (c), female, (d) 64 years and younger, (e) older than 64 years, (f) cvd, (g) resp, (h) inf, (i) uri, (j) ment hospitalization with a binary foehn threshold value of 72 which corresponds to 6 h of full foehn. The green line shows the temperature hospitallization association when foehn was present, the yellow line when foehn was abscent.*


```{r results_subgroups_interaction, echo=FALSE, fig.width=9, fig.height=9, fig.align='center'}
# subgroup analysis interaction plot

for (i in 1:length(groups_id[1:9])) {

  colvar  =  groups_id[i+1]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))
  
  # model with and without foehn
  mod_modif   <- gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction on the modifier
  pred_modif   <- crosspred(modif_rev, mod_modif, cen = 15.5, by = 0.1)
                   
  plot(pred_modif,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)
  text(-14, 2.4, labels = paste0("(", letters[i], ")"), pos = 2)
  
  }

```

*sFigure 7: Interaction between foehn wind and temperature with 95% confidence interval for (a) male, (b) female, (c) 64 years and younger, (d) older than 64, (e) circulatory, (f) respiratory, (g) infectious, (h) genitourinary, (i) mental hospitalizations.*
