---
title: "the direct effect of foehn winds on hospitalizations"
author: "Tino Schneidewind"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r data, echo=FALSE, warning=FALSE, message=FALSE}
library(dlnm);library(splines);library(ggplot2);library(viridis);library(gnm); library(dplyr);library(knitr);library(kableExtra);library(webshot2); library(magick)

data = read.csv() 

# index to include only stratum that have hospitalization counts
data$stratum_dow = as.factor(data$stratum_dow)
ind_dow = tapply(data$all, data$stratum_dow, sum)

# mmt function
source("../functions/findmin.R")

```

<br>

This Markdown document presents and summarizes the findings of an analysis concerning the following research question.

<br>

#### *Results for RQ1+RQ3: are hospitalizations independently affected by foehn winds in Switzerland? are certain subgroups (hospitalization cause/sex/age) a differentiating factor of this effect?*

<br>

*Table 1: Response ration for all-cause hospitalizations and subgroups at foehn wind value of 72 which corresponds to 6 h of full foehn. *

```{r resultss, echo=FALSE}

#### crossbasis temperature and foehn widns
cb.temp <- crossbasis(data$temperature,
                      lag=21,
                      argvar=list(fun="ns", knots = quantile(data$temperature, c(.5,.9), na.rm=TRUE)),
                      arglag=list(fun="ns", knots = logknots(21,3)),
                      group = data$station)

cb.foehn <- crossbasis(data$foehn_wind,
                       lag = 3,
                       argvar = list(fun="lin"),
                       arglag = list(fun="integer"),
                       group = data$station)


groups_id = colnames(data)[c(3,9,10,24,25, 13, 14, 11, 15, 12)]

table_estimates = data.frame(all = rep(NA, 2),
                             mal = rep(NA, 2),
                             fem = rep(NA, 2),
                             y64 = rep(NA, 2),
                             o64  = rep(NA, 2),
                             cvd = rep(NA, 2),
                             resp = rep(NA, 2),
                             inf = rep(NA, 2),
                             uri = rep(NA, 2),
                             ment = rep(NA, 2),
                             row.names = c("Model 1", "Model 2"))

# Model 1 + 2
for (i in 1:length(groups_id)) {

  colvar = groups_id[i]
  
  # formula, only foehn wind
  formula1 <- as.formula(paste(colvar, "~ cb.foehn"))

  # Model 1
  model_1 <- gnm(formula1, 
                 data = data,  
                 family=quasipoisson(), 
                 eliminate=stratum_dow, 
                 subset=ind_dow>0)

  # prediction
  pred_model_1 <- crosspred(cb.foehn, model_1, at=0:288, cumul=FALSE, cen = 0)

  # extract prediction for value 72 with confidence interaval and save it
  table_estimates[1,i] = paste0( sprintf("%.3f",round(pred_model_1$allRRfit["72"],digits=3)), 
                                         " [", 
                                         sprintf("%.3f",round(pred_model_1$allRRlow["72"], digits=3)),
                                         "-", 
                                         sprintf("%.3f",round(pred_model_1$allRRhigh["72"], digits = 3)), "]")
  
  
  
  # formula, foehn wind + temperature
  formula2 <- as.formula(paste(colvar, "~ cb.foehn + cb.temp"))

  # Model 2
  model_2 <- gnm(formula2, 
                 data = data,  
                 family=quasipoisson(), 
                 eliminate=stratum_dow, 
                 subset=ind_dow>0)

  # prediction
  pred_model_2 <- crosspred(cb.foehn, mod_nm2, at=0:288, cumul=FALSE, cen = 0)

  # extract prediction for value 72 with confidence intervals and save it
  table_estimates[2,i] = paste0( sprintf("%.3f",round(pred_model_2$allRRfit["72"],digits = 3)),
                                 " [", 
                                 sprintf("%.3f",round(pred_model_2$allRRlow["72"],digits=3)),
                                 "-", 
                                 sprintf("%.3f",round(pred_model_2$allRRhigh["72"], digits = 3)), "]")
  


}

# rotate table
table_estimates_transposed = t(table_estimates)

# finished table
kable(table_estimates_transposed,
      # digits = 3,
      col.names = c("Model 1", "Model 2"),
      align = rep("r", ncol(table_estimates_transposed) )) %>%
   kable_styling(full_width = FALSE) 

```






